---
title: "Predicting Airbnb Prices in Lisbon"
author: "Zsombor Hegedus & Fanni Kiss"
date: '2021 02 01 '
output: 
  html_document:
    code_folding: hide
---

## Executive summary

Our goal is to help our client price the rental fee for short time use of their small and mid-size apartments in Lisbon. In order to do so we built price predictor models based on a recent date: 20th Dec, 2020 using publicly available data scraped from Airbnb on [Inside Airbnb](http://insideairbnb.com/get-the-data.html). Our features captured information about the host, the number and quality of previous guest experiences, details of the accomodation and ultimately price, which was our target variable. We built several machine learning models such as OLS, Lasso, CART, GBM but in the end a random forest model was chosen as the best out of the lot. We evaluated the models with 5-fold cross validation both on a work and holdout set with RMSE being our loss function where the random forest was superior. We believe that our model is robust and can be of great help for our client both in case they want to advertise their accomodations with separate rooms or whole apartments as well. 

## Airbnb data 

In order to be able to build our models we did a heavy cleaning on the data downloaded from Inside Airbnb. We were informed that the client owns real estate in the heart of Lisbon only and that they own apartments that can accomodate 2-6 people only, so we narrowed our scope accordingly. We got rid of observations that do not belong to Lisbon or were not apartments (we also included condominimums) or apartment rooms for 2-6 people. Whether a property is a full apartment or a room is going to be very important furthermore in this analysis, because we want to show the client what options they have were they to decide to rent out not only full apartments but separate rooms as well. Furthermore, we excluded observations where the price was more than EUR1000/night as those were considered to be extreme values that are either erroneous or very much outside of the client's expectation. We aimed to keep as many variables as possible and dropped missing values only in rare cases, so we used imputation when that was necessary. The cleaning steps are documented in our [github repo](https://github.com/zsomborh/airbnb_lisbon/blob/main/code/clean_airbnb.R).

Overall we included variables that could be grouped into the following categories: 

- **Base Variables**: Such were the number of bedrooms, bathrooms, accommodates, the type of property which was either apartmant or room and the minimum nights one has to stay in the accommodation.
- **Host Information**: These variables were of the host of the apartmant such as the number of years they were hosts on Airbnb, how many listed apartmants they have, whether their identity is verified or if they are superhosts.
- **Reviews**: This category aims to capture the kind of feedback the accommodation received from the guests so far - these are the number of reviews, the average rating it received from clients from 1-10 and the amount of rating they receive in a month on average. 
- **Amenities**: These capture the extras that an accomodation can offer e.g.: whether it has a kitchen, coffee machine, or has locks on the door - from modeling perspective all such tags that were used more than 100 times were converted to dummies (e.g if accommodation has shampoo tag, it will receive a 1, and 0 otherwise). We decided not to keep them all as some tags appeared very infrequently and such low variation in a predictor would just throw off the model as opposed to improving it. 


```{r, warning=FALSE, message=FALSE}
rm(list = ls())

library(tidyverse)
library(data.table)
library(rattle)
library(caret)
library(ranger)
library(Hmisc)
library(knitr)
library(kableExtra)
library(xtable)
library(moments)
library(cowplot)
library(ggpubr)

#source(paste0(w_dir,'/code/helper.R'))
source('https://raw.githubusercontent.com/zsomborh/airbnb_lisbon/main/code/helper.R')

w_dir = 'C:/Users/T450s/Desktop/programming/git/airbnb_lisbon'
df <- read_csv('https://raw.githubusercontent.com/zsomborh/airbnb_lisbon/main/data/clean/cleaned_airbnb.csv')
```

Our target variable is price, a right skewed variable with a long right tail. On average in our sample apartmant costed almost twice as much as rooms, which is perfectly understandable given that the freedom of a whole apartment is something that people are willing to pay more. In total we have a bit more than 11k observations in which the number of apartments is almost four times more than the number of rooms. 

```{r table 1, echo = FALSE , results = "asis", warning = FALSE, message = FALSE }

price_sum <- df %>% group_by(property_type) %>% summarise(
    mean     = mean(price),
    median   = median(price),
    std      = sd(price),
    iq_range = IQR(price), 
    min      = min(price),
    max      = max(price),
    skew     = skewness(price),
    numObs   = sum( !is.na( price ) ) )

knitr::kable(price_sum, caption= 'Descriptive statistics of price for rooms and apartmants', digits = c(0,rep(2,ncol(price_sum)-2),0)) 
```

```{r fig 1, fig.width=10,fig.height=4, fig.cap='Distribution of apartment and room prices 20th Dec, 2020', echo = FALSE , results = "asis", warning = FALSE, message = FALSE }
p1 <- ggplot( data = df[df$property_type == 'apartment',] , aes( x = price ) ) +
    geom_histogram( aes(y = ..density..) , alpha = 1, binwidth = 30, color = 'black', fill = 'navyblue') +
    geom_density( aes(y = ..density..) , alpha = .5 , bw = 50, color = 'black', fill='steelblue') +
    labs(x='Price distribution of apartments',y='Density') + theme_minimal() 

p2 <- ggplot( data = df[df$property_type == 'room',] , aes( x = price ) ) +
    geom_histogram( aes(y = ..density..) , alpha = 1, binwidth = 30, color = 'black', fill = 'purple4') +
    geom_density( aes(y = ..density..) , alpha = .5 , bw = 50, color = 'black', fill='darkorchid2') +
    labs(x='Price distribution of rooms',y='Density') + theme_minimal() 

ggarrange(p1, p2, nrow = 1 )
```

We also created boxplots to further highlight the differnces in price for certain categories. We could see from the density plots but also in the first boxplot the nature of the difference in average price for rooms and apartments. Another important but trivial observation we can conclude from the second boxplot is that  


```{r fig 2, fig.width=10,fig.height=8, fig.cap='Boxplot of price by property type and number of accomodates', echo = FALSE , results = "asis", warning = FALSE, message = FALSE }

plotdf <- df %>% mutate(
    host_cat = ifelse(host_since <=3, '0-3 yrs', ifelse(host_since <= 7, '4-7 yrs','7+yrs')),
    score_cat = ifelse(review_scores_rating <= 80, '0-80', 
                       ifelse(review_scores_rating <= 85, '81-85', 
                              ifelse( review_scores_rating <= 90, '86-90',
                                      ifelse(review_scores_rating <= 95, '91-95', '96-100')))),
    min_cat = ifelse(minimum_nights ==1, '1 night',
                     ifelse(minimum_nights == 2, '2 nights', 
                        ifelse(minimum_nights == 3, '3 nights', '4+ nights')))) %>% 
    select(bedrooms,property_type, accommodates, host_cat,score_cat, price, min_cat)

p1 <- ggplot(plotdf, aes(x = factor(min_cat), y = price,
                        fill = factor(property_type), color=factor(property_type))) +
    geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
    scale_color_manual( name = '', values = c('black', 'black')) +
    scale_fill_manual( name = '', values = c('navyblue', 'purple4')) +
    stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
    labs(x = "Minimum nights",y = "Price (EUR)")+
    theme_minimal() +
    theme(legend.position = c(0.15,0.8))+
    scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 250), breaks = seq(0,250, 50))

p2 <- ggplot(plotdf, aes(x = factor(accommodates), y = price,
                        fill = factor(property_type), color=factor(property_type))) +
    geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
    scale_color_manual( name = '', values = c('black', 'black')) +
    scale_fill_manual( name = '', values = c('navyblue', 'purple4')) +
    stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
    labs(x = "Accomodates (Persons)",y = "Price (EUR)")+
    theme_minimal() +
    theme(legend.position = c(0.15,0.8))+
    scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 250), breaks = seq(0,250, 50))

p3 <- ggplot(plotdf, aes(x = factor(host_cat), y = price,
                        fill = factor(property_type), color=factor(property_type))) +
    geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
    scale_color_manual( name = '', values = c('black', 'black')) +
    scale_fill_manual( name = '', values = c('navyblue', 'purple4')) +
    stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
    labs(x = "How long host is present in Airbnb",y = "Price (EUR)")+
    theme_minimal() +
    theme(legend.position = c(0.15,0.8))+
    scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 200), breaks = seq(0,200, 50))

p4 <- ggplot(plotdf, aes(x = factor(score_cat), y = price,
                        fill = factor(property_type), color=factor(property_type))) +
    geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
    scale_color_manual( name = '', values = c('black', 'black')) +
    scale_fill_manual( name = '', values = c('navyblue', 'purple4')) +
    stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
    labs(x = "Review scores",y = "Price (EUR)")+
    theme_minimal() +
    theme(legend.position = c(0.15,0.8))+
    scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 200), breaks = seq(0,200, 50))


ggarrange(p1,p2,p3,p4, nrow = 2, ncol = 2)
```


## Data preparation

```{r, warning=FALSE, message=FALSE}

# some data prep

df <- df %>% mutate(
  property_type = as.factor(property_type),
  ln_price  = log(price)
)

colnames(df) <- tolower(colnames(df))


#variable grouping 
target_var = 'price'

basic_vars <- c('property_type', 'accommodates', 'bedrooms', 'beds', 'minimum_nights', 'bathrooms')

host_info <- c('host_since', 'host_is_superhost', 'host_listings_count', 'host_identity_verified')

reviews <- c('number_of_reviews', 'review_scores_rating','reviews_per_month', 'flag_review_scores_rating')

amenities <- colnames(df)[!colnames(df) %in% c('id',target_var, basic_vars, host_info,host_info,reviews)]
```

## Data exploration

```{r, warning=FALSE, message=FALSE}
# EDA

df %>%
  group_by(property_type) %>%
  dplyr::summarize(mean_price = mean(price, na.rm=TRUE))

Hmisc::describe(df$price)
```

```{r, warning=FALSE, message=FALSE}
g3a <- ggplot(data=df, aes(x=price)) +
  geom_histogram(aes(x = price), binwidth = 30, color = 'black', fill = 'navyblue', alpha = .8) +
  labs(x = "Price (EUR)",y = "Count")+
  theme_minimal() 
g3a
```

```{r, warning=FALSE, message=FALSE}
g3b <- ggplot(data=df, aes(x=ln_price)) +
  geom_histogram(aes(x = ln_price), binwidth = .2, color = 'black', fill = 'navyblue', alpha = .8) +
  labs(x = "Ln Price (EUR)",y = "Count")+
  theme_minimal() 
g3b
```

```{r, warning=FALSE, message=FALSE}
g4 <- ggplot(data = df, aes(x = property_type, y = price)) +
  stat_boxplot(aes(group = property_type), geom = "errorbar", width = 0.3,
               #color = c(color[2],color[1], color[3]), 
               size = 0.5, na.rm=T)+
  geom_boxplot(aes(group = property_type),
               #color = c(color[2],color[1], color[3]), fill = c(color[2],color[1], color[3]),
               size = 0.5, width = 0.6, alpha = 0.3, na.rm=T, outlier.shape = NA) +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,200), breaks = seq(0,200,50)) +
  labs(x = "Property_type",y = "Price (EUR)")+
  theme_minimal()


g4
```

```{r, warning=FALSE, message=FALSE}
g5 <- ggplot(df, aes(x = factor(accommodates), y = price,
                     fill = factor(property_type), color=factor(property_type))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  scale_color_manual( name = '', values = c('black', 'black')) +
  scale_fill_manual( name = '', values = c('navyblue', 'purple4')) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodates (Persons)",y = "Price (EUR)")+
  theme_minimal() +
  theme(legend.position = c(0.15,0.8))+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 250), breaks = seq(0,250, 50))
g5
```

```{r, warning=FALSE, message=FALSE}
# a few plots to check interactions
p1 <- price_diff_by_variables2(df, "property_type", "coffee_maker", "Property type", 'Coffee Maker')
p2 <- price_diff_by_variables2(df, "property_type", "wifi", "Property type", 'Wifi')
p3 <- price_diff_by_variables2(df, "property_type", "dishwasher", "Property type", 'Dishwasher') # looks good
p4 <- price_diff_by_variables2(df, "property_type", "elevator", "Property type", 'Elevator')
p5 <- price_diff_by_variables2(df, "property_type", "lock_on_bedroom_door", "Property type", 'Lock on bedroom')
p6 <- price_diff_by_variables2(df, "property_type", "microwave", "Property type", 'Microwave')
p7 <- price_diff_by_variables2(df, "property_type", "kitchen", "Property type", 'Kitchen') # looks good
p8 <- price_diff_by_variables2(df, "property_type", "lockbox", "Property type", 'LockBox') # looks good 
p9<- price_diff_by_variables2(df, "property_type", 'luggage_dropoff_allowed', "Property type",'Luggage dropoff')
p10<- price_diff_by_variables2(df, "property_type", 'patio_or_balcony', "Property type", 'Balcony avaialable') # looks good
p11<- price_diff_by_variables2(df, "property_type", 'refrigerator', "Property type", 'Fridge')
p12<- price_diff_by_variables2(df, "property_type", 'self_checkin', "Property type", 'Self Check-in') # looks good 

interaction_plot <- plot_grid(p1, p2, p3, p4, p5, p6, p7 ,p8, p9, p10, p11,p12, nrow=3, ncol=4)
interaction_plot

```

## Feature engineering

```{r, warning=FALSE, message=FALSE}
# Feature/target engineering -----------------------------------------------------

#Create log transformations on continuous variables where skewness is above .5 or below -5 - we  consider those to be skewed  

cont_vars <- c('minimum_nights', 'bedrooms','beds','bathrooms', 'host_since', 'host_listings_count', 'review_scores_rating', 'reviews_per_month', 'number_of_reviews')

df <- get_lns(df,cont_vars,0.5)

ln_vars <- colnames(df)[!colnames(df) %in% c(basic_vars,host_info,reviews,amenities,'id', 'price')]
ln_vars

# Create interaction variables - a few handpicked from EDA phase and all amenities + property type 
interactions1 <- c('property_type * dishwasher',
                   'property_type * kitchen',
                   'property_type * lockbox',
                   'property_type * patio_or_balcony',
                   'property_type * self_checkin'
)

interactions2 <- c(paste0('property_type * (',paste(amenities, collapse = '+'),')'))

# Transform outcome variable 

df <- df %>% mutate(
  ln_price = log(price)
)

# correct for ln 

df <- df %>% mutate(
  ln_minimum_nights =       ifelse(is.infinite(ln_minimum_nights),0,ln_minimum_nights),
  ln_bedrooms =             ifelse(is.infinite(ln_bedrooms),0,ln_bedrooms),
  ln_beds=                  ifelse(is.infinite(ln_beds),0,ln_beds),
  ln_bathrooms =            ifelse(is.infinite(ln_bathrooms),0,ln_bathrooms),
  ln_host_listings_count =  ifelse(is.infinite(ln_host_listings_count),0,ln_host_listings_count),
  ln_review_scores_rating = ifelse(is.infinite(ln_review_scores_rating),0,ln_review_scores_rating),
  ln_reviews_per_month =    ifelse(is.infinite(ln_reviews_per_month),0,ln_reviews_per_month),
  ln_number_of_reviews =    ifelse(is.infinite(ln_number_of_reviews),0,ln_number_of_reviews),
  ln_price =                ifelse(is.infinite(ln_price),0,ln_price)
)

for (i in c('ln_price',ln_vars)){
  print(paste0(i,'- no. infs is: ',sum(is.infinite(unlist(df[,i]))), ', no. of NAs is: ', sum(is.na(unlist(df[,i]))), ', no.of NaNs is: ', sum(is.nan(unlist(df[,i])))) )
}

sum(is.nan(df$ln_bathrooms))
#create predictors sets

predictors_1 <- c(basic_vars)
predictors_2 <- c(basic_vars, host_info, reviews, amenities)

transformed <- substring(ln_vars,4)

predictors_transformed_small <- c(predictors_2[!predictors_2 %in% transformed],interactions1,ln_vars)
predictors_transformed_big <- c(predictors_2[!predictors_2 %in% transformed],interactions2,ln_vars)

```


# Models

### 1. OLS

```{r, warning=FALSE, message=FALSE}
# Modeling with Lasso + OLS ----------------------------------------------------------------

#  create holdout set 
set.seed(7)

train_indices <- as.integer(createDataPartition(df$price, p = 0.75, list = FALSE))
df_train <- df[train_indices, ]
df_holdout <- df[-train_indices, ]


# train control is 5 fold cross validation
train_control <- trainControl(method = "cv",
                              number = 5,
                              verboseIter = FALSE)

# Starting off with simple OLSs 

set.seed(7)
system.time({
  ols_model1 <- train(
    formula(paste0("price ~", paste0(predictors_1, collapse = " + "))),
    data = df_train,
    method = "lm",
    trControl = train_control
  )
})
set.seed(7)
system.time({
  ols_model2 <- train(
    formula(paste0("price ~", paste0(predictors_2, collapse = " + "))),
    data = df_train,
    method = "lm",
    trControl = train_control
  )
})

set.seed(7)
system.time({
  ols_model3 <- train(
    formula(paste0("price ~", paste0(predictors_transformed_small, collapse = " + "))),
    data = df_train,
    method = "lm",
    trControl = train_control
  )
})



ols_model_coeffs1 <-  ols_model1$finalModel$coefficients
ols_model_coeffs_df1 <- data.frame(
  "variable" = names(ols_model_coeffs1),
  "ols_coefficient" = ols_model_coeffs1
) %>%
  mutate(variable = gsub("`","",variable))


ols_model_coeffs2 <-  ols_model2$finalModel$coefficients
ols_model_coeffs_df2 <- data.frame(
  "variable" = names(ols_model_coeffs2),
  "ols_coefficient" = ols_model_coeffs2
) %>%
  mutate(variable = gsub("`","",variable))


ols_model_coeffs3 <-  ols_model3$finalModel$coefficients
ols_model_coeffs_df3 <- data.frame(
  "variable" = names(ols_model_coeffs3),
  "ols_coefficient" = ols_model_coeffs3
) %>%
  mutate(variable = gsub("`","",variable))
```

### 2. LASSO

```{r, warning=FALSE, message=FALSE}
# Using Lasso

set.seed(7)
system.time({
  lasso_model <- train(
    formula(paste0("price ~", paste0(predictors_2, collapse = " + "))),  
    data = df_train,
    method = "glmnet",
    preProcess = c("center", "scale"),
    tuneGrid =  expand.grid("alpha" = 1, "lambda" = seq(0.01, 1, by = 0.01)),
    trControl = train_control
  )
})

set.seed(7)
system.time({
  lasso_model2 <- train(
    formula(paste0("price ~", paste0(predictors_transformed_big, collapse = " + "))), 
    data = df_train,
    method = "glmnet",
    preProcess = c("center", "scale"),
    tuneGrid =  expand.grid("alpha" = 1, "lambda" = seq(0.01, 1, by = 0.01)),
    trControl = train_control
  )
})


lasso_coeffs <- coef(
  lasso_model$finalModel,
  lasso_model$bestTune$lambda) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>%
  rename(lasso_coefficient = `1`) 

lasso_coeffs_non_null <- lasso_coeffs[!lasso_coeffs$lasso_coefficient == 0,]

lasso_coeffs2 <- coef(
  lasso_model2$finalModel,
  lasso_model2$bestTune$lambda) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>%
  rename(lasso_coefficient = `1`) 

lasso_coeffs_non_null2 <- lasso_coeffs2[!lasso_coeffs2$lasso_coefficient == 0,]

# TODO put everything in one 
regression_coeffs <- merge(ols_model_coeffs_df1,ols_model_coeffs_df2,ols_model_coeffs_df3, lasso_coeffs_non_null,lasso_coeffs_non_null2, by = "variable", all=TRUE)

# Check OLS, Lasso performance: 

temp_models <-
  list("OLS1" = ols_model1,
       "OLS2" = ols_model2,
       "OLS3" = ols_model3,
       "LASSO1 (model w/ few interactions)" = lasso_model,
       "LASSO2 (model w/ all interactions)" = lasso_model2)

result_temp <- resamples(temp_models) %>% summary()

result_temp
```

```{r, warning=FALSE, message=FALSE}
result_rmse <- imap(temp_models, ~{
  mean(result_temp$values[[paste0(.y,"~RMSE")]])
}) %>% unlist() %>% as.data.frame() %>%
  rename("CV RMSE" = ".")

result_holdout <- map(temp_models, ~{
  RMSE(predict(.x, newdata = df_holdout), df_holdout[["price"]])
}) %>% unlist() %>% as.data.frame() %>%
  rename("Holdout RMSE" = ".")

result_temp
result_rmse
result_holdout
```

### 3. Random forest

```{r, warning=FALSE, message=FALSE}
# insert random forest model
```

### 4. Gradient boosting machine

```{r, warning=FALSE, message=FALSE}
# insert GBM model
```

## Prediction

## Model selection

```{r, warning=FALSE, message=FALSE}
# Final models


final_models <-
  list("OLS3" = ols_model3,
       "LASSO (model w/ all interactions)" = lasso_model2,
       "CART" = cart_model,
       "Random forest" = rf_model,
       "GBM" = gbm_model)

results <- resamples(final_models) %>% summary()
results
```

```{r, warning=FALSE, message=FALSE}
#Evaluate models

#kable(x = result_4, format = "latex", digits = 3, booktabs=TRUE, linesep = "") %>%
#  cat(.,file= paste0(output,"horse_race_of_models_cv_rmse.tex"))

#Evaluate on holdout set 

result_5 <- map(final_models, ~{
  RMSE(predict(.x, newdata = df_holdout), df_holdout[["price"]])
}) %>% unlist() %>% as.data.frame() %>%
  rename("Holdout RMSE" = ".")

result_5
```

# Analytics

## Model diagnostics

### Variable importance plot

The TOP10 most important variables are shown on the plot below. 

```{r, warning=FALSE, message=FALSE}
# variable importance data frame
rf_model_var_imp <- importance(rf_model$finalModel)/1000
rf_model_var_imp_df <-
  data.frame(varname = names(rf_model_var_imp),imp = rf_model_var_imp) %>%
  arrange(desc(imp)) %>%
  mutate(imp_percentage = imp/sum(imp))

rf_model_var_imp_df
# variable importance plot - full plot
plot(varImp(rf_model))

# variable importance plot - top 10 only
rf_model_var_imp_plot_b <- ggplot(rf_model_var_imp_df[1:10,], aes(x=reorder(varname, imp), y=imp_percentage)) +
  geom_point(colour="navyblue", size=1) +
  geom_segment(aes(x=varname,xend=varname,y=0,yend=imp_percentage), colour="navyblue", size=0.75) +
  ylab("Importance (Percent)") +
  xlab("Variable Name") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=4), axis.text.y = element_text(size=4),
        axis.title.x = element_text(size=4), axis.title.y = element_text(size=4))
rf_model_var_imp_plot_b
```

### Grouped variable importance

We made four different groups from the variables, which were the following: 
* basic_vars: 'property_type', 'accommodates', 'bedrooms', 'beds', 'minimum_nights', 'bathrooms'
* host_info: 'host_since', 'host_is_superhost', 'host_listings_count', 'host_identity_verified'
* reviews: 'number_of_reviews', 'review_scores_rating','reviews_per_month', 'flag_review_scores_rating'
* amenities: group off the most important mentioned amenities

The graph below shows the importance of each group in the random forest model. The amenities together has 49% of importance in the random forest model.

```{r, warning=FALSE, message=FALSE}
rf_model_var_imp <- importance(rf_model$finalModel)/1000
rf_model_var_imp_df <-
  data.frame(varname = names(rf_model_var_imp),imp = rf_model_var_imp) %>%
  arrange(desc(imp)) %>%
  mutate(imp_percentage = imp/sum(imp))

rf_model_var_imp_df[rf_model_var_imp_df$varname %in% amenities,]

# grouped variable importance 2
# basic variables, host variables, review variables and amenities are grouped --> TO THE R MARKDOWN

rf_model_var_imp_df <- rf_model_var_imp_df %>% mutate(
  group2 = ifelse(varname %in% amenities, 'amenities',
                 ifelse(varname %in% basic_vars, 'basic_vars',
                        ifelse(varname %in% reviews, 'reviews', 'host_info'))))

rf_model_var_imp_grouped2 <- rf_model_var_imp_df %>%  group_by(group2) %>% summarise(group_imp_sum = sum(imp_percentage)) %>%  arrange(desc(group_imp_sum))

rf_model_var_imp_grouped2_plot <-
  ggplot(rf_model_var_imp_grouped2, aes(x=reorder(group2, group_imp_sum), y=group_imp_sum)) +
  geom_point(color="navyblue", size=1) +
  geom_segment(aes(x=group2,xend=group2,y=0,yend=group_imp_sum), color="navyblue", size=0.7) +
  ylab("Importance (Percent)") +   xlab("Variable Name") +
  coord_flip() +
  # expand=c(0,0),
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=4), axis.text.y = element_text(size=4),
        axis.title.x = element_text(size=4), axis.title.y = element_text(size=4))
rf_model_var_imp_grouped2_plot
```

If we unfold the groups, the difference between the importance of amenities and the rest of the variables are even larger. 

```{r, warning=FALSE, message=FALSE}
# grouped variable importance 3
# amenities are grouped --> TO THE R MARKDOWN

rf_model_var_imp_df <- rf_model_var_imp_df %>% mutate(
  group3 = ifelse(varname %in% amenities, 'amenities', varname))
                  
rf_model_var_imp_grouped3 <- rf_model_var_imp_df %>%  group_by(group3) %>% summarise(group_imp_sum = sum(imp_percentage)) %>%  arrange(desc(group_imp_sum))

rf_model_var_imp_grouped3_plot <-
  ggplot(rf_model_var_imp_grouped3, aes(x=reorder(group3, group_imp_sum), y=group_imp_sum)) +
  geom_point(color="navyblue", size=1) +
  geom_segment(aes(x=group3,xend=group3,y=0,yend=group_imp_sum), color="navyblue", size=0.7) +
  ylab("Importance (Percent)") +   xlab("Variable Name") +
  coord_flip() +
  # expand=c(0,0),
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=4), axis.text.y = element_text(size=4),
        axis.title.x = element_text(size=4), axis.title.y = element_text(size=4))
rf_model_var_imp_grouped3_plot
```

### Partial dependence plot

If we increase the number of accommodates, the price also increases in a fairly linear way.

```{r, warning=FALSE, message=FALSE}
pdp_n_acc <- pdp::partial(rf_model, pred.var = "accommodates", pred.grid = distinct_(df_holdout, "accommodates"), train = df_train)
pdp_n_acc_plot <- pdp_n_acc %>%
  autoplot( ) +
  geom_point(colour="navyblue", size=2) +
  geom_line(colour="navyblue", size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  scale_x_continuous(limit=c(1,7), breaks=seq(1,7,1))+
  theme_bw()
pdp_n_acc_plot
```

Comparing an apartment and a room, we can see that an apartment is predicted to have higher price than a room.

```{r, warning=FALSE, message=FALSE}
pdp_n_roomtype <- pdp::partial(rf_model, pred.var = "property_type", pred.grid = distinct_(df_holdout, "property_type"), train = df_train)
pdp_n_roomtype_plot <- pdp_n_roomtype %>%
  autoplot( ) +
  geom_point(color="navyblue", size=2) +
  ylab("Predicted price") +
  xlab("Room type") +
  scale_y_continuous(limits=c(60,120), breaks=seq(60,120, by=10)) +
  theme_bw()
pdp_n_roomtype_plot
```

### Subsample performance

The performance of the model in subsamples

```{r, warning=FALSE, message=FALSE}
# Subsample performance: 
# how well the rf_model performs on different subsets - calculating RMSE and the mean price per each subgroup

# holdout set
df_holdout_w_prediction <- df_holdout %>%
  mutate(predicted_price = predict(rf_model, newdata = df_holdout))


######### create nice summary table of heterogeneity
a <- df_holdout_w_prediction %>%
  mutate(is_low_size = ifelse(accommodates <= 3, "small apt", "large apt")) %>%
  group_by(is_low_size) %>%
  dplyr::summarise(
    rmse = RMSE(predicted_price, price),
    mean_price = mean(price),
    rmse_norm = RMSE(predicted_price, price) / mean(price)
  )




b <- df_holdout_w_prediction %>%
  filter(property_type %in% c("apartment", "room")) %>%
  group_by(property_type) %>%
  dplyr::summarise(
    rmse = RMSE(predicted_price, price),
    mean_price = mean(price),
    rmse_norm = rmse / mean_price
  )


c <- df_holdout_w_prediction %>%
  dplyr::summarise(
    rmse = RMSE(predicted_price, price),
    mean_price = mean(price),
    rmse_norm = RMSE(predicted_price, price) / mean(price)
  )

# Save output - RESULT_3 shows error
colnames(a) <- c("", "RMSE", "Mean price", "RMSE/price")
colnames(b) <- c("", "RMSE", "Mean price", "RMSE/price")
c <- cbind("All",c)
colnames(c) <- c("", "RMSE", "Mean price", "RMSE/price")

line1 <- c("Size of accomodation", "", "", "")
line2 <- c("Property type", "", "", "")
line3 <- c("All",  "", "", "")

result_3 <- rbind(line1, a, line2, b, line3, c) %>%
  transform(RMSE = as.numeric(RMSE), `Mean price` = as.numeric(`Mean price`),
            `RMSE/price` = as.numeric(`RMSE/price`))
```